\documentclass[usenames,dvipsnames,12pt]{report}
\usepackage[dvipsnames]{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subfig}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{minted}

\usetikzlibrary{
    calc,trees,positioning,arrows,chains,shapes.geometric,%
    decorations.pathreplacing,decorations.pathmorphing,shapes,%
    matrix,shapes.symbols
}
\tikzset{
    block/.style={rectangle, rounded corners, minimum height=3em, draw=black, text centered ,text width=5em},
    line/.style={->, thick,shorten >=1.5pt},
    decoration={brace},
    tuborg/.style={decorate},
    tubnode/.style={midway, right=2pt},
}

\begin{document}

\include{portada}

\tableofcontents

\listoffigures

\chapter{Introduction}

The goal of this document is to provide insight on how to implement a RSA cryptosystem
using \textbf{Python 3}. The system will be composed of two diferent modules named
\texttt{RSA} and \texttt{Key}. Containing the RSA class and the RSA's Key class respectively.\\

The system's architecture will be the following:

\begin{figure}[H]
    \centering
    \begin{tikzpicture}
        [node distance=1.5cm]

        \node[block] (rsa) {RSA.py};
        \node[block,right=of rsa] (key) {Key.py};
        \node[block, below=of rsa] (main) {main.py};

        \draw[line] (rsa) -- (key);
        \draw[line] (main) -- (rsa);

        \draw[dashed] (-2, -1.35) -- (10, -1.35);

        \draw[tuborg, decoration={brace}] let \p1=(key.north), \p2=(key.south) in
            ($(5.5, \y1)$) -- ($(5.5, \y2)$) node[tubnode] {RSA cryptosystem};

    \end{tikzpicture}

    % Caption and Label
    \caption{System architecture}
    \label{fig:system}
\end{figure}

The system is going to be built using a library pattern and in fact it will allow any program
to use it's contents to encrypt / decrypt messages. Consider RSA is also a slow algorithm compared
to symmetric algorithms.

\section{Rivest–Shamir–Adleman and public key cryptography}

RSA (Rivest–Shamir–Adleman) is a cryptosystem that was one of the first public-key based system.
That means that RSA have a public key and a private key used to share information between a sender
and a receiver.

\begin{figure}[H]
    \centering
    \begin{tikzpicture}
        [node distance=1cm]

        \node (sender) {Sender};
        \node[block, below=of sender] (encrypt) {Encrypt};
        \node[right=of sender] (channel) {Unsecure Channel};
        \node[right=of channel] (receiver) {Receiver};
        \node[block, below=of receiver] (decrypt) {Decrypt};

        \draw[line] (sender) -- (encrypt);
        \draw[line] (encrypt) -| (channel);
        \draw[line] (channel) -- (receiver);
        \draw[line] (receiver) -- (decrypt);

    \end{tikzpicture}

    % Caption and Label
    \caption{Transmission scheme}
    \label{fig:scheme}
\end{figure}

That said, the sender must first encrypt the message using the public key of the receiver.
Once the message is encrypted, it can be shared in a public channel that can be insecure by the pressence of
evasdroppers. The message can only be decrypted by the receiver by using it's private key. This can be achieved since
the public and the private keys share a mathematical relation. To further explain this, the following explanation
corresponds to the RSA cryptosystem.

\subsection{Key Generation}

To generate a RSA key you need to follow the following steps:

\begin{itemize}
    \item Choose two prime numbers $p$ and $q$.
        Those numbers should be randomly chosen and should have a similar length in bits.
    \item Calculate $n = p \cdot q$. The result $n$ will be used as the modulus for both
        keys (private, public) and it's length in bits designates the \textbf{key length}.
    \item Calculate $\Phi = (p - 1) \cdot (q - 1)$. Euler's properties are used to calculate $\Phi$.
    \item Calculate $e$ given $e < \Phi$ and $e$ to be coprime with $\Phi$.
    \item Determine $d$ given $e \cdot d \equiv 1 \mod \Phi$ (Modular multiplicative inverse)
\end{itemize}

The public key will be: \texttt{(n, e)} and the private key will be \texttt{(n, d)}

\subsection{Encryption}

Both, the sender and the receiver should share their public key in the public channel (it does not matter if anyone know it).
To demostrate an ecryption by using RSA let's consider the public key of the receiver as: \texttt{(n, e)} and it's
private key \texttt{(n, d)}.

If a sender wants to send the receiver a message it needs to encrypt the message by using the following mathematical formula:

\begin{center}
    $c \equiv m^e \mod n$
\end{center}

Consider $c$ to be the cipher text generated by the encryption.

\subsection{Decryption}

To decrypt, the same formula applies, but this time changing the $e$ with $d$ and $m$ with $c$.

\begin{center}
    $p \equiv c^d \mod n$ \\~\\
    $p \equiv (m^e)^d \mod n$ \\~\\
    $p \equiv m^{e \cdot d} \mod n \quad \rightarrow \quad p \equiv m^1 \mod n$ \\~\\
    $\boxed{p \equiv m \mod n}$
\end{center}

\chapter{Implementation}

This chapter will include the implementation details of how to implement the given RSA Cryptosystem
using \textbf{Python 3}. The implementation, as mentioned earlier, will be done as a library containing
two modules \texttt{Key.py} and \texttt{RSA.py}

\section{RSA Key}

The RSA Key module (\texttt{Key.py}) is the one responsible of the key creation for the RSA cryptosystem.\\

The Key class will be used inside the RSA class to determine the current used key. Keep in mind that
the Key is able to automatically execute the steps mentioned in the key generation process given
two integers $p$ and $q$. This is done in the constructor and further saved into the class attributes.\\

The class atributes and methods are defined below. The method names contains the attributes needed to work.
They also have an brief explanation of what they do.

\begin{center}
    \begin{tabular}{ | l | p{6cm} |}
    \hline
    p, q, n, phi, e, d & The basic storage attributes for the RSA key as defined in the introduction.\\ \hline
    \_\_init\_\_(self, p, q) & Basic constructor to generate a key given $p$ and $q$.\\ \hline
    \_\_generate\_e\_and\_d(self) & Method to generate an e and d given $\Phi$ and $e$.\\ \hline
    \_\_generate\_e(self, phi) & Generate an $e$ based on $\Phi$.\\ \hline
    \_\_extended\_euclidean\_algorithm(self, a, b) & The Extended Euclidean Algorithm given $a = e$ and $b = \Phi$ to compute $d$ considering $d \cdot e \equiv \mod \Phi$.\\ \hline
    length(self) & Returnns the length of the key.\\ \hline
    public(self) & Returns the public key.\\ \hline
    private(self) & Returns the private key.\\ \hline
    print(self) & Prints the whole key (Public + Private).\\
    \hline
    \end{tabular}
\end{center}

\inputminted[linenos, breaklines=true]{python}{../Python/RSA/Key.py}

\section{RSA Cryptosystem}

The RSA Class module (\texttt{RSA.py}) is the one responsible of the RSA cryptosystem.\\

The RSA class will be responsible of generating keys, encrypting and decrypting messages.
The key generation will be dispatched to the Key class, responsible of this things. The
RSA will only store and use the key.

The class atributes and methods are defined below. The method names contains the attributes needed to work.
They also have an brief explanation of what they do.

\begin{center}
    \begin{tabular}{ | l | p{6cm} |}
    \hline
    key & Stores the RSA key class used to encrypt / decrypt messages.\\ \hline
    \_\_fast\_modular\_exponentiation(self, a, z, n) & Method to calculate $x \equiv a^z \mod n$.\\ \hline
    generate\_key(self, p, q) & Method to generate a key given $p$ and $q$\\ \hline
    encrypt(self, plain\_text) & Encrypts a message (Might be a number or a string)\\ \hline
    decrypt(self, cipher\_text) & Decrypts an encrypted number or sequence of numbers (string).\\
    \hline
    \end{tabular}
\end{center}

\inputminted[linenos, breaklines=true]{python}{../Python/RSA/RSA.py}

\chapter{Library Usage}

To use the library you simply need to import it as a regular \texttt{Python 3} library and use the RSA class
to access the given public methods. That said, the following example may ilustrate a sample usage case.

\inputminted[linenos, breaklines=true]{python}{../Python/main.py}

\chapter{References}

Khan Academy. (2019). \textit{Fast modular exponentiation}.
[online] Available at:
\url{https://www.khanacademy.org/computing/computer-science/cryptography/modarithmetic/a/fast-modular-exponentiation}
[Accessed 12 Jan. 2019].\\

Loyola University Chicago. (2019). \textit{Extended Euclidean Algorithm}. 1st ed.
[ebook] Boston, Massachusetts: Dr. Andrew Harrington, p.2. Available at:
\url{https://anh.cs.luc.edu/331/notes/xgcd.pdf}
[Accessed 12 Jan. 2019].\\

Wikipedia. (2019). \textit{RSA (cryptosystem)}.
[online] Available at:
\url{https://en.wikipedia.org/wiki/RSA_(cryptosystem)}
[Accessed 12 Jan. 2019].\\

Python. (2019). \textit{Python 3.7.2 Documentation}.
[online] Available at:
\url{https://docs.python.org/3/}
[Accessed 12 Jan. 2019].

\end{document}
